\hypertarget{course-schedule}{%
\section{Course Schedule}\label{course-schedule}}

\textbf{Subject To Change}. All changes will be announced in-class,
through email, and on the \href{https://lms.cofc.edu}{OAKS} course page.

\vspace{0.10in}

\noindent \textbf{All readings listed under \emph{Readings} below are
required}. The readings listed below will be available on
\href{https://lms.cofc.edu}{OAKS} as a pdf or as an external link.
Readings for each week can be found under Content. I recommend doing the
readings in the order listed below.

\SetDate[06/01/2022]

\hypertarget{part-i-foundations}{%
\subsection{Part I: Foundations}\label{part-i-foundations}}

\week \textbf{Introduction and Course Overview}

\begin{itemize}

\item
  \emph{Readings (skim after class)}:

  \begin{itemize}
  
  \item
    \textbf{Nowlin}, Chapter 2: \emph{A Framework for Environmental
    Policymaking}, from Nowlin, Matthew C. 2019. \emph{Environmental
    Policymaking in an Era of Climate Change}. New York, NY: Routledge.
  \end{itemize}
\end{itemize}

\week \textbf{Environmental Risks, Law, and Enforcement}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Lazarus}, Chapter 2: \emph{The Implications of Ecological
    Injury for Environmental Protection Law}, from Lazarus, Richard J.
    2004. \emph{The Making of Environmental Law}. Chicago, IL:
    University of Chicago Press.
  \item
    Shaffer, Rachel M. 2021. ``Environmental Health Risk Assessment in
    the Federal Government: A Visual Overview and a Renewed Call for
    Coordination.'' \emph{Environmental Science \& Technology} 55(16):
    10923--27.
  \item
    \textbf{Salzman and Thompson}, Chapter 2, pgs. 31-49: \emph{Four
    Analytical Frameworks}
  \item
    \textbf{Salzman and Thompson}, Chapter 4: \emph{Enforcement}, from
    Salzman, James, and Barton H. Thompson Jr.~2014. \emph{Environmental
    Law and Policy}. 4th ed.~St.~Paul, MN: Foundation Press.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 1}
  \end{itemize}
\end{itemize}

\week \textbf{Actors, Administrative Law, and the Environment}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Rowell and van Zeben}, Chapter 2: \emph{Key Actors} and
    Chapter 3: \emph{Types of Laws}, from Rowell, Arden, and Josephine
    van Zeben. 2021. \emph{A Guide to U.S. Environmental Law}. Oakland,
    CA: University of California Press.
  \item
    \textbf{Rosenbloom}, Chapter 1: \emph{What is Administrative Law}?,
    from Rosenbloom, David H. 2015. \emph{Administrative Law for Public
    Managers}. Boulder, CO: Westview Press.
  \item
    \textbf{SKIM}: Tatel, David S. 2010. ``The Administrative Process
    and the Rule of Environmental Law.'' \emph{Harvard Environmental Law
    Review} 34: 1--8.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 2}
  \end{itemize}
\end{itemize}

\week \textbf{Environmental Regulation and Rulemaking}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Dudley and Brito}, Chapter 2: \emph{Theories of Regulation:
    Why Do We Regulate}?, from Dudley, Susan E., and Jerry Brito. 2012.
    \emph{Regulation: A Primer}. 2nd Edition. Arlington, VA: Mercatus
    Center at George Mason University.
  \item
    \textbf{Rowell and van Zeben}, Chapter 4: \emph{Regulatory
    Instruments}, from Rowell, Arden, and Josephine van Zeben. 2021.
    \emph{A Guide to U.S. Environmental Law}. Oakland, CA: University of
    California Press.
  \item
    \textbf{Rosenbloom}, Chapter 3: \emph{Administrative Rulemaking},
    from Rosenbloom, David H. 2015. \emph{Administrative Law for Public
    Managers}. Boulder, CO: Westview Press.
  \item
    \href{https://www.epa.gov/laws-regulations/regulations}{Regulations
    (EPA Website)}
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 3}
  \end{itemize}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{part-ii-legal-and-regulatory-applications-to-pollution}{%
\subsection{Part II: Legal and Regulatory Applications to
Pollution}\label{part-ii-legal-and-regulatory-applications-to-pollution}}

\week \textbf{Air Pollution: I}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \href{https://www.epa.gov/laws-regulations/summary-clean-air-act}{Summary
    of the Clean Air Act}
  \item
    \textbf{Bearden et al.}, \emph{Clean Air Act}, pgs. 3-24, from
    Bearden, David M. et al.~2013. \emph{Environmental Laws: Summaries
    of Major Statutes Administered by the Environmental Protection
    Agency}. Washington D.C.: Congressional Research Service.
  \item
    Boyd, William. 2019. ``The Clean Air Act's National Ambient Air
    Quality Standards: A Case Study of Durability and Flexibility in
    Program Design and Implementation.'' In \emph{Lessons from the Clean
    Air Act: Building Durability and Adaptability into U.S. Climate and
    Energy Policy}, eds.~Ann Carlson and Dallas Burtraw. Cambridge
    University Press, 15--56.
  \item
    \textbf{SKIM}: Liu, Jiawen et al.~2021. ``Disparities in Air
    Pollution Exposure in the United States by Race/Ethnicity and
    Income, 1990--2010.'' \emph{Environmental Health Perspectives}
    129(12).
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 4 due}
  \end{itemize}
\end{itemize}

\week \textbf{Air Pollution: II}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    Rabe, Barry G. 2019. ``Leveraged Federalism and the Clean Air Act:
    The Case of Vehicle Emissions Control.'' In Lessons from the Clean
    Air Act: Building Durability and Adaptability into US Climate and
    Energy Policy, eds.~Ann Carlson and Dallas Burtraw. Cambridge:
    Cambridge University Press, 113--58.
  \item
    Patashnik, Eric M. 2019. ``The Clean Air Act's Use of Market
    Mechanisms.'' In Lessons from the Clean Air Act: Building Durability
    and Adaptability into US Climate and Energy Policy, eds.~Ann Carlson
    and Dallas Burtraw. Cambridge: Cambridge University Press, 201--24.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 5 due}
  \end{itemize}
\end{itemize}

\week \textbf{Climate Change}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    Crystal, Howard M., Kassie Siegel, Maya Golden-Krasner, and Clare
    Lakewood. 2019. ``Returning to Clean Air Act Fundamentals: A Renewed
    Call to Regulate Greenhouse Gases Under the National Ambient Air
    Quality Standards (NAAQS) Program.'' \emph{The Georgetown
    Environmental Law Review} 31(2): 233--84.
  \item
    \href{https://www.vox.com/platform/amp/2021/11/3/22758188/climate-change-epa-clean-power-plan-supreme-court}{A
    new Supreme Court case could gut the government's power to fight
    climate change}
  \item
    \href{https://www.scotusblog.com/case-files/cases/west-virginia-v-environmental-protection-agency/}{SCOTUSblog:
    West Virginia v. Environmental Protection Agency}
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 6 due}
  \end{itemize}
\end{itemize}

\week \textbf{Water Pollution}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \href{https://www.epa.gov/laws-regulations/summary-clean-water-act}{Summary
    of the Clean Water Act}
  \item
    \textbf{Bearden et al.}, \emph{Clean Water Act}, pgs. 25-33, from
    Bearden, David M. et al.~2013. \emph{Environmental Laws: Summaries
    of Major Statutes Administered by the Environmental Protection
    Agency}. Washington D.C.: Congressional Research Service.
  \item
    \textbf{Salzman and Thompson}, Chapter 7: \emph{Water Pollution},
    from Salzman, James, and Barton H. Thompson Jr.~2014.
    \emph{Environmental Law and Policy}. 4th ed.~St.~Paul, MN:
    Foundation Press.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{‌Reflection paper 7 due}
  \end{itemize}
\end{itemize}

\week \textbf{Spring Break}

\week \textbf{Drinking Water}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \href{https://www.epa.gov/sdwa}{Safe Drinking Water Act (SDWA) EPA
    Website}
  \item
    Tiemann, Mary. 2017. Safe Drinking Water Act (SDWA): A Summary of
    the Act and Its Major Requirements. Washington, D.C.:
    \emph{Congressional Research Service}.
  \item
    Switzer, David, and Manuel P. Teodoro. 2017. ``The Color of Drinking
    Water: Class, Race, Ethnicity, and Safe Drinking Water Act
    Compliance.'' \emph{Journal AWWA} 109(9): 40--45.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{‌Reflection paper 8 due}
  \end{itemize}
\end{itemize}

\week \textbf{Toxic Substances and Waste Management}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Salzman and Thompson}, Chapter 8: \emph{Regulating Toxic
    Substances}
  \item
    \textbf{Salzman and Thompson}, Chapter 9: \emph{Waste Management},
    from Salzman, James, and Barton H. Thompson Jr.~2014.
    \emph{Environmental Law and Policy}. 4th ed.~St.~Paul, MN:
    Foundation Press.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{‌Reflection paper 9 due}
  \end{itemize}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{part-iii-legal-and-regulatory-applications-to-ecosystem-services}{%
\subsection{Part III: Legal and Regulatory Applications to Ecosystem
Services}\label{part-iii-legal-and-regulatory-applications-to-ecosystem-services}}

\week \textbf{Public Trust, Wetlands, and Waters of the United States}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Salzman and Thompson}, Chapter 10: \emph{Wetlands,
    Endangered Species, and the Public Trust}, \textbf{pgs. 271-287},
    from Salzman, James, and Barton H. Thompson Jr.~2014.
    \emph{Environmental Law and Policy}. 4th ed.~St.~Paul, MN:
    Foundation Press.
  \item
    Finney, Brad. 2018. ``One WOTUS, Two POTUS: The Clean Water Act and
    the Economic Impact.'' \emph{Tennessee Law Review} 86(4): 895--964.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{‌Reflection paper 10 due}
  \end{itemize}
\end{itemize}

\week \textbf{Public Lands and Endangered Species}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \textbf{Andrews}, Chapter 13: \emph{Public Lands and Wildlife
    Conservation}, from Andrews, Richard N. L. 2020. \emph{Managing the
    Environment, Managing Ourselves: A History of American Environmental
    Policy}. Third Edition. New Haven, CT: Yale University Press
  \item
    \textbf{Salzman and Thompson}, Chapter 10: \emph{Wetlands,
    Endangered Species, and the Public Trust}, \textbf{pgs. 287-310},
    from Salzman, James, and Barton H. Thompson Jr.~2014.
    \emph{Environmental Law and Policy}. 4th ed.~St.~Paul, MN:
    Foundation Press.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 11}
  \end{itemize}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{part-iv-environmental-impact-statements}{%
\subsection{Part IV: Environmental Impact
Statements}\label{part-iv-environmental-impact-statements}}

\week \textbf{The National Environmental Policy Act}

\begin{itemize}

\item
  \emph{Readings}:

  \begin{itemize}
  
  \item
    \href{https://www.epa.gov/laws-regulations/summary-national-environmental-policy-act}{Summary
    of the National Environmental Policy Act}
  \item
    \textbf{Salzman and Thompson}, Chapter 12: \emph{The National
    Environmental Policy Act}, from Salzman, James, and Barton H.
    Thompson Jr.~2014. \emph{Environmental Law and Policy}. 4th
    ed.~St.~Paul, MN: Foundation Press.
  \item
    Scott, Tyler Andrew, Nicola Ulibarri, and Omar Perez Figueroa. 2020.
    ``NEPA and National Trends in Federal Infrastructure Siting in the
    United States.'' \emph{Review of Policy Research} 37(5): 605--33.
  \end{itemize}
\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Reflection paper 12}
  \end{itemize}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\week \textbf{Conclusion}

\begin{itemize}

\item
  \emph{Assignments}:

  \begin{itemize}
  
  \item
    \textbf{Paper}
  \item
    \textbf{Presentation}
  \end{itemize}
\end{itemize}
